---
title: "ACDS1010 Data in R I Examples"
author: "Sean Hellingman"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data in R Examples


Some examples using the class() function:

```{r class_function}
?class() #help

O1 <- FALSE
O2 <- 'FALSE'
O3 <- NA
O4 <- 14.5

class(O1) #logical
class(O2) #character
class(O3) #logical
class(O4) #numeric
```



## Example 1 using class() from slides





Some examples using the is.[classOfData]() function:

```{r is.class_function}
?is.numeric() #help

O1 <- FALSE
O2 <- 'FALSE'
O3 <- NA
O4 <- 14.5

is.numeric(O1) #logical
is.numeric(O2) #character
is.numeric(O3) #logical
is.numeric(O4) #numeric


is.character(O1) #logical
is.character(O2) #character
is.character(O3) #logical
is.character(O4) #numeric
  
  
is.logical(O1) #logical
is.logical(O2) #character
is.logical(O3) #logical
is.logical(O4) #numeric  
  
```






Some examples using converting functions:

```{r class_function}
?as.character() #help

O1 <- FALSE
O2 <- 'FALSE'
O3 <- NA
O4 <- 14.5

#Characters#
as.character(O1) #logical
as.character(O2) #character
as.character(O3) #logical
as.character(O4) #numeric

#Numeric#
as.numeric(O1) #logical
as.numeric(O2) #character ##Problem!!
as.numeric(O3) #logical
as.numeric(O4) #numeric
```




## Example 2 from slides:







This code is what you use to load packages and data:
```{r loading-packages}

library(dplyr) #Load the package
data("starwars") #load the data



```



## Example 3 from the slides






## Example 4 from the slides

```{r loading_CSVs}
?read.csv() #help

#data <- read.csv("filename.csv", header = TRUE) #Use this




```




The following code loads some American stock exchange symbols from Yahoo Finance:
```{r American-Symbols}
if(!require(quantmod)) install.packages("quantmod")
library(quantmod)

??quantmod #Package help
?getSymbols() #help


### vector of American Stock components
vecOfSymbs = c("AXP", "GE", "GS", "HD", "NKE")  


getSymbols(vecOfSymbs, source = "yahoo", from = "1995-01-03", to = "1999-12-31")


```


## Example 5 from slides






## Example 6 from slides

```{r Saving}

?save() #help
#save(GE,file = "GE.RData")

?write.csv() #help
#write.csv(GE,file = "GE.csv")

```

