---
title: "ACDS1010 R Markdown Examples"
author: "Student Name and Number"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


If you need to install the packages on your own:

```{r packages}


#install.packages("tinytex")
#tinytex::install_tinytex()


```




# Heading 1

## Heading 2

### Heading 3


**Bold**

*italic*

~~Here is a list:~~

- item 1
- item 2
  + sub-item 2
  + sub-item 3
- Superscript^2^
- Subscript~2~


<!-- This is a R markdown comment -->

Here is the [link](https://rstudio.github.io/cheatsheets/html/rmarkdown.html) to the R markdown cheat sheet.


The following code chunk called summary-stats loads the *mtcars* package from base R. Then it outputs a summary of the data. **ALL CODE CHUNKS MUST HAVE DIFFERENT NAMES**

```{r, summary-stats}
data("mtcars") #Imports mtcars data from base R
#?mtcars
summary(mtcars) #summary of the variables in mtcars
```


The following code chunk plots the horsepower by the number of cylinders from the cars data.
```{r, plot1}
plot(mtcars$cyl,mtcars$hp, ylab = 'Horsepower',xlab = 'Cylinders')
```


The following code chunk creates the same plot without rendering the code, resizes the plot, and adds a figure caption.


```{r, plot1-resized, echo=FALSE, fig.cap="Horesepower by Cylinders", fig.width=4, fig.height=3 }
plot(mtcars$cyl,mtcars$hp, ylab = 'Horsepower',xlab = 'Cylinders')
```

\newpage


*Note:* The two colon symbols specify which package a function is being called from.

The following code generates a table of the first ten observations from the first four columns of the *mtcars* dataset. The code also changes the column names.

```{r, table1}
library(knitr) #We need to call the knitr library

knitr::kable(head(mtcars[,0:4]),
col.names = c('MPG', 'Cylinders',
'Displacement', 'HP'))
```



# In Class Examples


## Example 2



## Example 3


In the R markdown file you have already created, do the following:

- Initialize a vector called Vector.1, that goes from 1 to 10.
- Initialize a vector called Vector.2 that goes from 50 to 5 by 5.
- In a separate code chunk, plot Vector.1 by Vector.2.
- In a separate code chunk, plot Vector.1 by Vector.2 but adjust the size to be 1 inch high and 2 inches wide.
- Add a caption to the smaller figure.

```{r, Vectors-ex3}



# Vector.2 <- seq(50, 5, by = -5)
# plot(x,y)
```




```{r, plot1-resized-example3, echo=FALSE, fig.cap="", fig.width=1, fig.height=1 }


```


## Example 4
In the R markdown file you have already created, do the following:

- Use data("iris") to load the iris dataset in base R.
- Create a table using kable() of the first 10 observations from this data.
- Rename the columns to something more suitable


```{r, table-ex4}
# library(knitr) #We need to call the knitr library
# 
# knitr::kable(head(mtcars[,0:4]),
# col.names = c('MPG', 'Cylinders',
# 'Displacement', 'HP'))



```






