---
title: "ADSC1010 Data Manipulation in R dplyr Examples"
author: "Sean Hellingman"
date: "2023-09-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
rm(list = ls()) #Remove all objects from the workspace


if(!require(dplyr)) install.packages("dplyr")
library(dplyr)



```



```{r select-cars}

data("mtcars")
?select() # help
colnames(mtcars)

New_Cars <- dplyr::select(mtcars,mpg,cyl,hp,wt)
head(New_Cars)


```

## Example 1 from slides


```{r select-example}

data("starwars")
colnames(starwars)




```



```{r filter-cars }
?filter() # help
summary(mtcars$qsec)


Fast_Cars <- filter(mtcars,qsec <= 17)
head(Fast_Cars)


```


## Example 2 from slides

```{r filter-example}



```







```{r mutate-cars}

#Kilometres per litre#
mtcars <- mutate(mtcars, kpl = mpg/2.352)
head(mtcars)
```



## Example 3 from slides

```{r mutate-example}
?mutate() #help
#starwars$mass/(starwars$height/100)^2
colnames(starwars)



```



Arranging by two variables really only works if there are multiple observations for some groups.
The order of the variables that you arrange by matters.

```{r arrange-cars}

mtcars <- arrange(mtcars,cyl, -hp)
head(mtcars)

```


## Example 4 from slides

```{r arrange-example}
?arrange() #help


```




Summarize

```{r summarize-cars}
#mean and median hp#

cars_summary <- summarize(mtcars,mean_hp = mean(hp), median_hp = median(hp))
cars_summary

```


## Example 5 from slides

```{r summarize-example}
?summarize() #help
na.omit()



```





#### Sequential operations

Average displacement (disp) of cars with over 109 hp and maximum number of cylinders. 

```{r sequential-cars}


#Intermediary

#over 109 hp
cars_109 <- filter(mtcars, hp > 109)
#max cylinders
max_cyl_109_hp <- filter(cars_109, cyl == max(cyl))
#final value
Av_disp <- summarize(max_cyl_109_hp,av_disp = mean(disp))
Av_disp


#anonymous
Av_disp_1 <- summarise(
  filter(
    filter(
      mtcars, hp > 109
    ),
    cyl == max(cyl)
  ),
  av_disp = mean(disp)
)

```





## Example 5 from slides

```{r seq-operations-example}

#Intermediary




#anonymous



```



Using the pipe operator %>% 

```{r pipe-cars-example}

Av_disp_2 <- mtcars %>% 
  filter(hp > 109) %>% 
  filter(cyl == max(cyl)) %>% 
  summarise(avg_disp = mean(disp))



```





Using the *group_by()* function.
```{r cars-group_by}


Cars_Averages <- mtcars %>% 
  group_by(cyl) %>% 
  summarise(
    avg_hp_cyl = mean(hp),
    avg_mpg_cyl = mean(mpg)
  )

Cars_Averages

```





## Example 6 from slides

```{r group_by-example}




```









# Joining Data 



```{r data for example}

set.seed(12)

# Create a small dataset of students (only interested in these students)
small_students <- data.frame(
  StudentID = c(1, 2, 3, 4),
  Name = c("Alice", "Bob", "Charlie", "Sean")
)

# Create a large dataset of courses with more students
large_courses <- data.frame(
  CourseID = c(101, 102, 103, 104, 105),
  CourseName = c("Math", "Science", "History", "English", "Art"),
  StudentID = sample(1:15, 15)
)

Small_Grades <- left_join(small_students,large_courses, by = "StudentID")
Small_Grades
```





## Example 7 from slides

```{r import-data}




```

